---
title: "HSM_9_35"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HSM_9_35}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The 2010 Highway Safety Manual (HSM) Section 9-35 Site Specific Empirical Bayes (EB) Analysis is included in version 1.0 of HSM package. The vignette provides instructions to use the section 9-35 "HSM_9_35" and "spf" functions.

```{r setup}
#install.packages("devtools")	
library(devtools)
```

Install package via GitHub.
```{r}
#install_github("https://github.com/cn838/HSM.git")
library(HSM)
```

# The spf function
Predicted crashes can be computed using the spf function. 
```{r}
spf(
  AADTMAJ=1000,
  AADTMIN=100,
  L=NULL,
  Base_Condition="HSM-RUR2-4ST-KABCO",
  Provide_Overdispersion_Factor=FALSE,
  Segment=FALSE
)
```

The overdispersion parameter (k) is provided by specifying TRUE within the provide overdispersion factor parameter.  Please note that version 1.0 only includes intersection SPFs from the 2010 HSM found in the SPF_table.RData. 
```{r}
spf(
  AADTMAJ=1000,
  AADTMIN=100,
  L=NULL,
  Base_Condition="HSM-RUR2-4ST-KABCO",
  Provide_Overdispersion_Factor=TRUE,
  Segment=FALSE
)
```
# The HSM Section 9-35 function

The HSM_9_35 function is demonstrated within two examples.  The two examples demonstrate the accuracy and the use of the function with HSM example and user defined data, respectively. 

## A sample problem to illustrate the HSM_9_35 function with the EB Before/After Safety Effectiveness Evaluation Method

Excel friendly users may import data into Rstudio. It is recommended to use the X9_10_template with version 1.0. Take for example, section 9.10  saved as "X9_10_seg.RData" (pg. 9-17). The script supports a total of 20 years of data (e.g. 10 years before and after).

```{r}
str(X9_10_seg) 
```

Check the "Basic Input Data" using the "testthat" package.
```{r}
library(testthat)

testthat::test_that("Basic Input Data check", {
  expect_equal(sum(HSM::HSM_9_35(data = X9_10_seg, segment=TRUE, group = FALSE, group_base_by = "Past")[[9]]), 122)
  expect_equal(sum(HSM::HSM_9_35(data = X9_10_seg, segment=TRUE, group = FALSE, group_base_by = "Past")[["Observed Crashes in the After Period"]]), 30)
  print("Basic input data matches.")
})

```

The site specific results are obtained and cross verified with the HSM.
```{r}
testthat::test_that("values from Eq. 9A.1-11", {
  expect_equal(sum(HSM::HSM_9_35(data = X9_10_seg, segment=TRUE, group = FALSE, group_base_by = "Past" )[["Variance (9A.1-11)"]]), 11.161 )
  print("The cumulative sum of the Variance terms match the expected output")
})

```

Similarly, the "Estimation of the Precision of the Treatment Effectiveness" are verified.
```{r}
testthat::test_that("Precision tests", {expect_equal(HSM::HSM_9_35(data = X9_10_seg, segment=TRUE, group = TRUE, group_base_by = "Past")[["Standard Error of Safety Effectiveness (9A.1-13)"]], 13.882)
  expect_equal(HSM::HSM_9_35(data = X9_10_seg, segment=TRUE, group = TRUE, group_base_by = "Past")[["Test Statistic (Step 14)"]], 2.194)
  print("The percision outputs match the results from Steps 13 & 14")
  })
```

The site specific results table is now ready to be copied and pasted into a spreadsheet, for word processing. 
```{r}
HSM::HSM_9_35(data = X9_10_seg, segment=TRUE, group = FALSE, group_base_by = "Past") %>% knitr::kable(digits = 3, format.args = list(decimal.mark = ".", big.mark = ","), align= 'c', caption = "Example 9.10 Emperical Bayes Total Crash Pooled Results ")
```

Lastly, the effectiveness of the treatment effect or CMF is obtained, for word processing.
```{r}
HSM::HSM_9_35(data = X9_10_seg, segment=TRUE, group = TRUE, group_base_by = "Past") %>% knitr::kable(digits = 3, format.args = list(decimal.mark = ".", big.mark = ","), align= 'c', caption = "Example 9.10 Emperical Bayes Total Crash Pooled Results ")
```

## HSM_9_35 for Intersections
Entering data is performed by implementing base R functions (i.e. concatenate c() and data.frame()).  Please note that version 1.0 requires the column names to be exactly the same as contained in the "X9_10_template.RData". 
```{r}
data=data.frame(
Site_No.=c(1,2,3,4,5,6), # id number 
Yrs_Before=c(3,3,3,3,3,3), # years in the before period
Yrs_After=c(3,3,3,3,3,3), # years in the after period
Base_Past=c(
  "HSM-RUR2-3ST-KABCO",	#SPF associated with the intersection or treatment for KABCO or KABC crash types.
  "HSM-RUR2-4ST-KABCO", 
  "HSM-RUR2-4SG-KABCO",
  "HSM-RUR2-3ST-KABCO",	
  "HSM-RUR2-4ST-KABCO",
  "HSM-RUR2-4SG-KABCO"
),
Base_Current=c(
  "RUR2-4RA",
  "RUR2-4RA",
  "RUR2-4RA",
  "Urban Single Lane Roundabout",
  "RUR2-4RA",
  "RUR2-4RA"),
L=c(NA,NA,NA,NA,NA,NA), # omit for intersection analysis
CMF=c(1.1,0.90,0.98,1,1,1.0), #crash modification factor. User specifies 1.0 if none was used.
C=c(1.2,1.2,1.2,1.2,1.2,1.2), #calibration factor. User specifies 1.0 if none was used.

Before.AADT.Major_1=c(1100,1190,1100,1150,900,500), #traffic volumes on the major approach during year one for site one ect.
Before.AADT.Major_2=c(1000,900,1150,1150,900,550),
Before.AADT.Major_3=c(1000,950,1125,1150,900,550),
Before.AADT.Major_4=c(1000,950,1125,1150,900,550),
Before.AADT.Major_5=c(1050,900,1125,1150,900,550),
Before.AADT.Major_6=c(1050,900,1125,1150,900,550),

Before.AADT.Minor_1=c(100,90,1100,110,90,550), #traffic volumes for minor approaches
Before.AADT.Minor_2=c(100,90,1100,110,90,550),
Before.AADT.Minor_3=c(100,90,1100,110,90,550),
Before.AADT.Minor_4=c(100,90,1100,110,90,550),
Before.AADT.Minor_5=c(100,90,1100,110,90,550),
Before.AADT.Minor_6=c(100,90,1100,110,90,550),

After.AADT.Major_1=c(1000,1000,1000,1000,1000,1000), #major approach traffic volumes in the after period 
After.AADT.Major_2=c(1000,1000,1000,1000,1000,1000),
After.AADT.Major_3=c(1000,1000,1000,1000,1000,1000),
After.AADT.Major_4=c(1000,1000,1000,1000,1000,1000),
After.AADT.Major_5=c(1000,1000,1000,1000,1000,1000),
After.AADT.Major_6=c(1000,1000,1000,1000,1000,1000),

After.AADT.Minor_1=c(200,200,80,700,80,90),#minor approach traffic volumes in the after period 
After.AADT.Minor_2=c(200,200,80,700,80,90),
After.AADT.Minor_3=c(200,200,80,700,80,90),
After.AADT.Minor_4=c(200,200,80,700,80,90),
After.AADT.Minor_5=c(200,200,80,700,80,90),
After.AADT.Minor_6=c(200,200,80,700,80,90),

Before.Yr_1=c(0,2,3,1,1,3), # crashes in the before periods 1,2,3,..,6
Before.Yr_2=c(0,0,0,0,0,0),
Before.Yr_3=c(1,0,1,3,0,0),
Before.Yr_4=c(0,1,0,1,0,1),
Before.Yr_5=c(1,0,0,1,0,1),
Before.Yr_6=c(0,0,0,1,1,0),

After.Yr_1=c(0,0,0,0,0,1), # crashes in the after periods 1,2,3,..,6
After.Yr_2=c(0,0,0,0,0,0),
After.Yr_3=c(0,0,1,0,0,1),
After.Yr_4=c(0,1,0,0,0,0),
After.Yr_5=c(0,0,0,0,0,0),
After.Yr_6=c(1,0,0,1,0,0))
```
Specify the "group" parameter equal to "FALSE" to provide a  tibble that contains the site specific results. 
```{r}
HSM::HSM_9_35(data = data, segment=FALSE, group = FALSE)
```
Specify the "group" and "group_base_by" parameters equal to "TRUE", "Past" or "Current" to provide a tibble that contains the "Estimation of the Precision of the Treatment Effectiveness". The unbiased Odds Ratio is the Crash Modification Factor (CMF). Similar to the Comparison Group (CG) method, sites with zero crashes in the after periods will effect the precision of the estimate.  
```{r}
HSM::HSM_9_35(data = data, segment=FALSE, group = TRUE, group_base_by= "Past")
```

Specify the "group_base_by" parameter to obtain results aggregated by future or current base conditions.   
```{r}
HSM::HSM_9_35(data = data, segment=FALSE, group = TRUE, group_base_by = "Current") 
```

The results are copied and pasted into an spreadsheet, for word processing using "kable()".
```{r}
HSM::HSM_9_35(data = data, segment=FALSE, group = TRUE, group_base_by = "Current" ) %>% knitr::kable(digits = 3, format.args = list(decimal.mark = ".", big.mark = ","), align= 'c', caption = "Table 1: Estimation of the Precision of the Treatment Effectiveness")
```
Please note more flexibility will be provided with future revisions.
